Supporting information: Tail associations and their impact on extinction risk

### 2019-11-10 MainText.Rmd
formatting for journal

---
title: "Tail associations and their impact on extinction risk"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
date: ""
geometry: "left=1in,right=1in,top=1in,bottom=1in"
output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty

tables: True
link-citations: True
urlcolor : blue
indent : True

@@

\vspace{12pt}
\noindent \textbf{\textit{Keywords:}} extinction risk, extreme events, tail association, density-dependent model, copula, mathematical ecology

@@

<!--FIG :  summary : all nonlinear model results for local dispersion : extrisk vs. d-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=14.5 cm]{./Results/allmodel_ext_risk_local_disp_extreme_and_moderate_noise.pdf}
\caption{Extinction risk versus dispersal rate for different scenarios of tail association
for our six metapopulation models (Methods, Table \ref{tab_model_summary}). Dispersal was
local, see Fig. \ref{SM-fig_extrisk_allmodel_globaldisp} for global dispersal. Values of $r$
used are in the panel headers.
Extinction risks are for a 25-year time horizon. Confidence intervals for extinction 
risk are not plotted, to provide a clearer plot. But 
because 10000 simulations were
done for each extinction risk estimate, confidence intervals
were very narrow, as in Fig. \ref{fig_extrisk_agestr}. LT=left tail, RT=right 
tail. \label{fig_extrisk_allmodel_localdisp}}
\end{center}
\end{figure*}

>>>>

---
title: "Tail associations and their impact on extinction risk"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
date: ""
fontsize: 12pt
geometry: "left=1in,right=1in,top=1in,bottom=1in"
output: 
  pdf_document:
    number_sections: no
    keep_tex: yes
    fig_caption: yes
    includes:
      in_header: head_maintext.sty

mainfont: Times New Roman      
tables: True
link-citations: True
urlcolor : blue
indent : True

@@

\vspace{12pt}
\noindent \textbf{\textit{Keywords:}} copula; density-dependent model; extinction risk; extreme events; mathematical ecology; tail association.

@@

<!--FIG :  summary : all nonlinear model results for local dispersion : extrisk vs. d-->
\begin{figure*}[!h]
\begin{center}
\includegraphics[width=14 cm]{./Results/allmodel_ext_risk_local_disp_extreme_and_moderate_noise.pdf}
\caption{Extinction risk versus dispersal rate for different scenarios of tail association
for our six metapopulation models (Methods, Table \ref{tab_model_summary}). Dispersal was
local, see Fig. \ref{SM-fig_extrisk_allmodel_globaldisp} for global dispersal. Values of $r$
used are in the panel headers.
Extinction risks are for a 25-year time horizon. Confidence intervals for extinction 
risk are not plotted, to provide a clearer plot. But 
because 10000 simulations were
done for each extinction risk estimate, confidence intervals
were very narrow, as in Fig. \ref{fig_extrisk_agestr}. LT=left tail, RT=right 
tail. \label{fig_extrisk_allmodel_localdisp}}
\end{center}
\end{figure*}

### 2019-11-10 SuppMat.Rmd
formatting for journal

---
title: "Supporting information: Tail associations and their impact on extinction risk"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
date: ""
geometry: "left=1in,right=1in,top=1in,bottom=1in"

@@

tables: True
link-citations: True
urlcolor : blue
indent : True

@@

# Generating noise with asymmetric tail associations  \label{makenoise}

<!--# Algorithm to generate extreme tail association  \label{extr_tail_dep}-->
Bivariate environmental noise with extreme left-tail association (as in Fig. 
\ref{MT-fig_intro_cop}A) was generated using an algorithm originally presented 
by @Ghosh_copula (their Appendix S10), presented again here for convenience.
If $T$ time steps of noise were desired, step 1 was to generate a $T$ by $2$
matrix, $M$, of independent random numbers, each standard-normally distributed. 
Then, each row of $M$ was identified for modification randomly and
independently, with a 50\% chance. For those rows identified, every element
of the row was replaced by the negative of the absolute value of the first element
of the row, including the first element itself. For the other rows, every 
element in the row was replaced by its own absolute value. For noise
with extreme right-tail association (as in Fig. 
\ref{MT-fig_intro_cop}D), take the negative of the entire matrix. For
both types of noise, every entry was transformed by the cumulative distribution
function of a standard normal distribution, to make each column uniformly
distributed.
The Spearman correlation of the extreme left- or right-tail associated 
noise was `r srho_common`, up to sampling variation. A derivation
of this fact is presented below.

>>>>

---
title: "Supporting information: Tail associations and their impact on extinction risk"
author: "Shyamolina Ghosh, Lawrence W. Sheppard, Daniel C. Reuman"
date: ""
fontsize: 12pt
geometry: "left=1in,right=1in,top=1in,bottom=1in"

@@

mainfont: Times New Roman       
tables: True
link-citations: True
urlcolor : blue
indent : True

@@

# Generating noise with asymmetric tail associations\label{makenoise}

Bivariate environmental noise with extreme left-tail association (as in Fig. 
\ref{MT-fig_intro_cop}A) was generated using an algorithm originally presented 
by @Ghosh_copula (their Appendix S10), presented again here for convenience.
If $T$ time steps of noise were desired, step 1 was to generate a $T$ by $2$
matrix, $M$, of independent random numbers, each standard-normally distributed. 
Then, each row of $M$ was identified for modification randomly and
independently, with a 50\% chance. For those rows identified, every element
of the row was replaced by the negative of the absolute value of the first element
of the row, including the first element itself. For the other rows, every 
element in the row was replaced by its own absolute value. For noise
with extreme right-tail association (as in Fig. 
\ref{MT-fig_intro_cop}D), take the negative of the entire matrix. For
both types of noise, every entry was transformed by the cumulative distribution
function of a standard normal distribution, to make each column uniformly
distributed.
The Spearman correlation of the extreme left- or right-tail associated 
noise was `r srho_common`, up to sampling variation. A derivation
of this fact is presented below.
